Renewal and stability in populations structured by remaining years of life

### 2016-11-29 ThanoRepro/ThanoRepro.tex
starting to erase former lifetable transformation work. It distracts from the renewal aspect of the paper.

Each of these elements is structured by
age, $a$. The model can be approximated by observed data because
age is typically known. % The survival function, $\ell(a)$, derives
%from a lifetable, usually based on age-specific event-exposure mortality rates,
%which require counts of decedents and an estimate of the
%population exposed to risk in the same chronological age classes.
Age in the growth factor, $e^{-ra}$, pertains to the passage of
time more than it does to chronological age (this is synchronous with birth
cohorts). The product of this growth factor and $\ell(a)$ gives a
function proportional to the stable age structure, $c(a)$, and it is often
intuitive to imagine these two elements together as one for this reason. Female
fertility rates, $m(a)$, are based on daughters born to mothers in this limited
single-sex model. The functions $m(a)$ and $\ell(a)$
together intuitively make for net female fertility, and their age-wise product
is sometimes written as $\phi(a)$. This model lends itself to commonly collected data because age
is typically known, and can be asked.
% For the two essential rate componenents underlying the renewal
 % equation~\eqref{eq:lotka}, classification by chronological age is essentially given--- it is deterministic in the sense that time passed since birth is typically known by individuals. The kinds of errors in such classification are mostly due to innumeracy, rounding, exaggeration, and similar kinds of biases, but even in such cases, chronological age is fundamentally knowable.

@@

The basic transformation to years left starts with the well-known remaining
years density function, $f(y|a)$.
%For age zero, $f(y|0)$ is identical to $d(a)$, the lifetable deaths
% distribution (based on a unity radix).
%For ages higher than 0, simply condition on survival:
%\footnote{In the
% reliability literature, $f(y|a)$ is denoted by $ X - t | X > t $, where $X$ is the total lifespan and $t$ is the
%current chronological age attained.}
\begin{equation}
\label{eq:vaupel1}
f(y | a) = \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \ec
\end{equation}
where $\mu(a)$ is the force of mortality.
%A convenient discretization of \eqref{eq:vaupel1} is to simply work with the
%$d_a$ column of the lifetable (here with single ages):
%\begin{equation}
%\label{eq:dyadiscret}
%f_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \ec
%\end{equation}
%where $L_a$ is lifetable exposure.\footnote{Equation~\eqref{eq:dyadiscret}
%will incur some error depending on the method used to derive $L_a$. To remove
%this error, rescale over the $a$ margin so that each distribution conditioned
% on $a$ sums to 1. Working with $\ell_a$ rather than $L_a$ eliminates error and
%yields similar results, but is not necessarily better.} One produces in this
% way a probability density function over future death times for each chronological age, $a$. 
The $f(y|a)$- weighted average of $y$ for a given $a$ returns the remaining life
expectancy column of the lifetable, $e(a)$. $f(y|a)$ is also useful for
decomposing age-classified data, such as the
population pyramid, into remaining years classes. If $P(a)$ is an
age-classified population count, we can derive the years left version $P(y)$,
using\footnote{An identical definition is found in
\citet{brouard1989mouvements} and illustrated in \citet{brouard1986structure}.}

@@

Figure~\ref{fig:pop} shows this exercise for 2010 US data \citep{HMD},
assuming homogeneity and that the 2010 period mortality schedule is held fixed.
When looking from the age perspective (Figure~\ref{fig:USdecomp}), this
decomposition reveals projected years left heterogeneity within a stock of known age, but when viewed from the
years left perspective (Figure~\ref{fig:USrecomp}) it reveals age heterogeneity
and the projected overall years left profile--- the future death flow of the
current population, $P(y)$ under our assumptions.

% both figures created in R/Pyramids.R
\begin{figure}[ht!]
	\caption{2010 US population structure}
	\label{fig:pop}
	\begin{center}
	\begin{subfigure}{.45\textwidth}
		\caption{Age structure (years lived) decomposed by years left.}
		\label{fig:USdecomp}
		\includegraphics[scale=.6]{Figures/US2010Age.pdf}
	\end{subfigure}
	~
	\begin{subfigure}{.45\textwidth}
		\caption{Years left structure decomposed by years lived.}
		\label{fig:USrecomp}
		\includegraphics[scale=.6]{Figures/US2010Thano.pdf}
	\end{subfigure}
	\\
	\small{*Population and mortality data from the HMD.}
	\end{center}
\end{figure}

\citet{brouard1986structure} applied this method to highlight regional
differences in French age structure, at times combining this projective method
with historical stocks. \citet{brouard1989mouvements}, and \citet{vaupel2009life}, showed that in stationary populations the
marginal chronological and thanatological age structures are identical, and this
provides a limited basis for comparison between the two perspectives. 
%\citet{wachter2014essential} hints at, and
%\citet{vaupel2014stable} elaborates on how to make chronological and
%thanatological age structure commensurate under a constant growth rate, $r$.
%These results also follow from the following specification of a
%thanatological renewal model.

Lifetable-derived thanatological age structure is
probabilistic until after death in several senses. In this case, note that
counts from a given age class are distributed over a range of higher ages
according to a given mortality rate schedule; in this
case a period mortality pattern is used.
As such, Figure~\ref{fig:USrecomp} is \textit{projective} in nature, a
forward-looking glance at population attrition given the population stock and lifetable of a
particular moment, whereas Figure~\ref{fig:USdecomp} is \textit{reflective} in
nature, since a population's chronological age-structure is the fruit of past
demographic forces.

\subsection{Thanatological fertility rates.}
Any count
classified by chronological age can be reclassified as in
equation~\eqref{eq:transform} given a pertinent lifetable. To
derive thanatological fertility rates under these assumptions, $\gamma(y)$, apply \eqref{eq:transform} to
chronological age-classified birth counts, $B(a)$, to get $B(y)$ and again to
exposure-to-risk, $E(a)$, to get $E(y)$ and then divide in the usual way:

\begin{equation}
\gamma(y) = \frac{B(y}{E(y)} \ep
\end{equation}

Figure~\ref{fig:FySpaghetti}
represents the full variety in female thanatological period fertility rates that can be found for all years
of data that overlap in the HFD and HMD.\footnote{There are as of this writing
1933 population-years of overlap between the HMD and HFD, including a wide
variety of fertility and mortality combinations.} I show $\gamma(y)$ calculated in this way because
it enters into the renewal model here studied. Figure~\ref{fig:FxSpaghetti}
gives ASFR for the same populations and years as a more familiar reference, but
note that both the x and y axes are different. The shape of thanatological
fertility curves is speculative and due to the above-mentioned strong assumptions. Nonetheless it is best not to
include such rates into a model blindly, so we offer this schematic overview of
some of their characteristics.

Period thanatological fertility rates estimated in this way have a
characteristic shape. The distribution is wider than that of chronological
fertility rates. The range of post-childbearing lifespans stretches from 
zero, due e.g., to maternal mortality, to at least 100.\footnote{For instance,
Jeanne Calment lived for another 100 years after giving birth to her daughter.}
This spaghetti plot includes both fertility booms and busts, as well as some
mortality crises (1918, WWII). Some patterns in the data (not
highlighted in the figure) include that the left tail (a gauge of how
orphan-prone a population is) has tended to fall over time. All populations have shown an increase in both the
mean and mode of thanatological age at childbearing over time. Several
contemporary populations now show modes of over 60 years. Period thanatological total
fertility rates track standard period total fertility rates closely, but tend to be somewhat higher (not shown).
% both figures created in R/HFDspaghetti.R
\begin{figure}[h!]
	\caption{Chronological and thantological fertility rates, all 1933
	country-year combinations present in both the HFD and HMD.* Note different x
	and y scales.}
	\label{fig:Fxcompare}
	\begin{center}
	\makebox[\textwidth]{
	\begin{subfigure}{.45\textwidth}
		\caption{Chronological period fertility rates.}
		\label{fig:FxSpaghetti}
		\includegraphics[scale=.45]{Figures/FxSpaghettiDraft.png}
	\end{subfigure}
	~
	\begin{subfigure}{.45\textwidth}
		\caption{Thanatological period fertility rates.}
		\label{fig:FySpaghetti}
		\includegraphics[scale=.45]{Figures/FySpaghettiDraft.png}
	\end{subfigure}
	}
	\\
	\end{center}
	\begin{tiny}
	*AUT, $1951-2010$; BGR, $1947-2009$; BLR, $1964-2012$; CAN, $1921-2009$; 
	CHE, $1932-2011$; CHL, $1992-2005$; CZE, $1950-2011$; DEUTE, $1956-2011$; 
	DEUTNP, $1990-2011$; DEUTW, $1956-2011$; ESP, $1922-2012$; EST, $1959-2011$; 
	FIN, $1939-2009$; FRATNP, $1946-2013$; GBR\_NIR, $1974-2011$; GBR\_NP,
	$1974-2011$; GBR\_SCO, $1945-2011$; GBRTENW, $1938-2011$; HUN, $1950-2009$;
	IRL, $1955-2009$; ISL, $1963-2010$; JPN, $1947-2012$; LTU, $1959-2011$; NLD, $1950-2009$; 
	NOR, $1967-2009$; PRT, $1940-2012$; RUS, $1959-2010$; SVK, $1950-2009$; 
	SVN, $1983-2009$; SWE, $1891-2011$; TWN, $1976-2010$; UKR, $1959-2009$; 
	USA, $1933-2010$
	\end{tiny}
\end{figure}

In general, thanatological fertility rates will not be useful for purposes of
projection, unless their empirical regularity is
stronger than typical ASFR. This is a question I do not seek to answer
in the present treatment, but in any case such rates are not susceptible to
direct measurement.
There are alternative ways of defining thanatological fertility rates, for instance by transforming the birth and
population vector from a stable population. As proven in appendix~\ref{app:B},
thanatological fertility rates derived from a stable population imply a certain
symmetry with stable populations from the chronological model.

\FloatBarrier
\section*{The thanatological renewal model}
The renewal model that follows uses thanatological fertility rates, $\gamma(y)$,
the form of which is suggested in Figure~\ref{fig:FySpaghetti}, and we refer to
a homogenous, closed, unisex population based on female vital rates. Assuming constant vital rates, the births for the present year are given by 
\begin{align}
B(t) =& \int _0^\infty P(y,t)\gamma(y) \dd y = \int _0^\infty
P(a,t)m(a) \dd a \ec \intertext{where $a$ indexes chronological age, $y$ indexes
thanatological age, $P(a), P(y)$ are population counts, and $\gamma(y), m(a)$ are
exact fertility probabilities (rates), best imagined
as the fertility rates of females born to females. The thanatological
(first) integral can be broken down back in terms of chronological age:} =&
\int_{y=0}^\infty \int_{a=0}^\infty \gamma(y)P(a,t)f(y|a) \dd a \dd y \ec
\intertext{where $f(y|a)$ is defined per equation~\eqref{eq:vaupel1}, so
$d(a)/\ell(0)$.
We can relate the present population to past births with $P(a,t) = B(t-a)\ell(a)$:}
\label{eq:ergostep1}
=& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) B(t-a)d(a+y)\dd a \dd y \ep 
\intertext{Eventually strong ergodicity will assert itself, and $B(t)$ will
be related to $B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
intrinsic rate of growth:}
\label{eq:ergostep2}
=& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) B(t)e^{-ra}d(a+y)\dd a \dd y
\ep
\intertext{Divide out $B(t)$ to get back a familiar-looking renewal equation:}
\label{eq:thanoren}
1 =& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) d(a+y)e^{-ra}\dd a \dd y
\ep
\end{align}
Now compare this to Lotka's chronological formulation in
Equation~\eqref{eq:lotka},which can be expressed as well as:
\begin{equation}
 1 = \int_{a=0}^\infty \int_{y=0}^\infty m(a)d(a+y)e^{-ra}\dd y \dd a \ec
  \label{eq:lotka2}
\end{equation}
and note that \eqref{eq:lotka2} and \eqref{eq:thanoren} are really quite
similar, since $\int _{y=0}^\infty d(a+y)\dd a = l(a)$. For intuition, notice
that $l(a)$ is here decomposed into $d(a)$. Imagine a matrix of this
decomposition, where one margin is chronological age and the other margin is
thanatological age.
For the chronological case \eqref{eq:lotka}, we multiply chronological
age-specific fertility rates, $m(a)$, over the chronological age margin, and for the thanatological case we multiply $\gamma(y)$ over the thanatological age margin.

>>>>

Each of these elements is structured by
age, $a$. %The model can be approximated by observed data because
%age is typically known. % The survival function, $\ell(a)$, derives
%from a lifetable, usually based on age-specific event-exposure mortality rates,
%which require counts of decedents and an estimate of the
%population exposed to risk in the same chronological age classes.
Age in the growth factor, $e^{-ra}$, pertains to the passage of
time more than it does to chronological age (this is synchronous with birth
cohorts). The product of this growth factor and $\ell(a)$ gives a
function proportional to the stable age structure, $c(a)$, and it is often
intuitive to imagine these two elements together as one for this reason. Female
fertility rates, $m(a)$, are based on daughters born to mothers in this limited
single-sex model. The functions $m(a)$ and $\ell(a)$
together intuitively make for net female fertility, and their age-wise product
is sometimes written as $\phi(a)$. This model lends itself to commonly collected data because age
is typically known, and can be asked.
% For the two essential rate componenents underlying the renewal
 % equation~\eqref{eq:lotka}, classification by chronological age is essentially given--- it is deterministic in the sense that time passed since birth is typically known by individuals. The kinds of errors in such classification are mostly due to innumeracy, rounding, exaggeration, and similar kinds of biases, but even in such cases, chronological age is fundamentally knowable.

@@

The basic transformation to years left starts with the well-known remaining
years density function, $f(y|a)$.
%For age zero, $f(y|0)$ is identical to $d(a)$, the lifetable deaths
% distribution (based on a unity radix).
%For ages higher than 0, simply condition on survival:
%\footnote{In the
% reliability literature, $f(y|a)$ is denoted by $ X - t | X > t $, where $X$ is the total lifespan and $t$ is the
%current chronological age attained.}
\begin{equation}
\label{eq:vaupel1}
f(y | a) = \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \ec
\end{equation}
where $\mu(a)$ is the force of mortality.
%A convenient discretization of \eqref{eq:vaupel1} is to simply work with the
%$d_a$ column of the lifetable (here with single ages):
%\begin{equation}
%\label{eq:dyadiscret}
%f_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \ec
%\end{equation}
%where $L_a$ is lifetable exposure.\footnote{Equation~\eqref{eq:dyadiscret}
%will incur some error depending on the method used to derive $L_a$. To remove
%this error, rescale over the $a$ margin so that each distribution conditioned
% on $a$ sums to 1. Working with $\ell_a$ rather than $L_a$ eliminates error and
%yields similar results, but is not necessarily better.} One produces in this
% way a probability density function over future death times for each chronological age, $a$. 
The $f(y|a)$- weighted average of $y$ for a given $a$ returns the remaining life
expectancy column of the lifetable, $e(a)$. $f(y|a)$ is also useful for
decomposing age-classified data, such as the
population pyramid, into years-left classes. If $P(a)$ is an
age-classified population count, under fixed mortality we can infer the
years-left structure, $P(y)$, using\footnote{An identical definition is found in
\citet{brouard1989mouvements} and illustrated in \citet{brouard1986structure}.}

@@

%Figure~\ref{fig:pop} shows this exercise for 2010 US data \citep{HMD},
%assuming homogeneity and that the 2010 period mortality schedule is held fixed.
%When looking from the age perspective (Figure~\ref{fig:USdecomp}), this
%decomposition reveals projected years left heterogeneity within a stock of known age, but when viewed from the
%years left perspective (Figure~\ref{fig:USrecomp}) it reveals age heterogeneity
%and the projected overall years left profile--- the future death flow of the
%current population, $P(y)$ under our assumptions.

% both figures created in R/Pyramids.R
%\begin{figure}[ht!]
%	\caption{2010 US population structure}
%	\label{fig:pop}
%	\begin{center}
%	\begin{subfigure}{.45\textwidth}
%		\caption{Age structure (years lived) decomposed by years left.}
%		\label{fig:USdecomp}
%		\includegraphics[scale=.6]{Figures/US2010Age.pdf}
%	\end{subfigure}
%	~
%	\begin{subfigure}{.45\textwidth}
%		\caption{Years left structure decomposed by years lived.}
%		\label{fig:USrecomp}
%		\includegraphics[scale=.6]{Figures/US2010Thano.pdf}
%	\end{subfigure}
%	\\
%	\small{*Population and mortality data from the HMD.}
%	\end{center}
%\end{figure}

\citet{brouard1986structure} applied this method to highlight regional
differences in French age structure, at times combining this projective method
with historical stocks. \citet{brouard1989mouvements}, and \citet{vaupel2009life}, showed that in stationary populations the
marginal years-lived and years-left distributions are identical,
and this provides a limited basis for comparison between the two perspectives. 
%\citet{wachter2014essential} hints at, and
%\citet{vaupel2014stable} elaborates on how to make chronological and
%thanatological age structure commensurate under a constant growth rate, $r$.
%These results also follow from the following specification of a
%thanatological renewal model.

Lifetable-derived years-left structure is
probabilistic until after death in several senses. In this case, note that
counts from a given age class are distributed over a range of higher ages
according to a given mortality rate schedule; in this
case a period mortality pattern is used.
Redistribution as in \eqref{eq:transform} is \textit{projective}, a
forward-looking glance at population attrition given the population stock and lifetable of a
particular moment, whereas the classic pyramid is \textit{reflective}, since a
population's age-structure owes to past demographic forces.

Other age-structured quantities can be reclassified under the same assumptions.
If birth counts and population exposures are reclassified as in
\eqref{eq:transform}, their ratio yields like-classified fertility rates,
$\gamma{y}$.

%\subsection{Years-left fertility rates.}
%Any age-classified can be reclassified as in
%equation~\eqref{eq:transform} given a pertinent lifetable. To
%derive thanatological fertility rates under these assumptions, $\gamma(y)$, apply \eqref{eq:transform} to
%chronological age-classified birth counts, $B(a)$, to get $B(y)$ and again to
%exposure-to-risk, $E(a)$, to get $E(y)$ and then divide in the usual way:
%
%\begin{equation}
%\gamma(y) = \frac{B(y}{E(y)} \ep
%\end{equation}
%
%Figure~\ref{fig:FySpaghetti}
%represents the full variety in female thanatological period fertility rates that can be found for all years
%of data that overlap in the HFD and HMD.\footnote{There are as of this writing
%1933 population-years of overlap between the HMD and HFD, including a wide
%variety of fertility and mortality combinations.} I show $\gamma(y)$ calculated in this way because
%it enters into the renewal model here studied. Figure~\ref{fig:FxSpaghetti}
%gives ASFR for the same populations and years as a more familiar reference, but
%note that both the x and y axes are different. The shape of thanatological
%fertility curves is speculative and due to the above-mentioned strong assumptions. Nonetheless it is best not to
%include such rates into a model blindly, so we offer this schematic overview of
%%some of their characteristics.

%Period thanatological fertility rates estimated in this way have a
%characteristic shape. The distribution is wider than that of chronological
%fertility rates. The range of post-childbearing lifespans stretches from 
%zero, due e.g., to maternal mortality, to at least 100.\footnote{For instance,
%Jeanne Calment lived for another 100 years after giving birth to her daughter.}
%This spaghetti plot includes both fertility booms and busts, as well as some
%mortality crises (1918, WWII). Some patterns in the data (not
%highlighted in the figure) include that the left tail (a gauge of how
%orphan-prone a population is) has tended to fall over time. All populations have shown an increase in both the
%mean and mode of thanatological age at childbearing over time. Several
%contemporary populations now show modes of over 60 years. Period thanatological total
%fertility rates track standard period total fertility rates closely, but tend to be somewhat higher (not shown).
%% both figures created in R/HFDspaghetti.R
%\begin{figure}[h!]
%	\caption{Chronological and thantological fertility rates, all 1933
%	country-year combinations present in both the HFD and HMD.* Note different x
%	and y scales.}
%	\label{fig:Fxcompare}
%	\begin{center}
%	\makebox[\textwidth]{
%	\begin{subfigure}{.45\textwidth}
%		\caption{Chronological period fertility rates.}
%		\label{fig:FxSpaghetti}
%		\includegraphics[scale=.45]{Figures/FxSpaghettiDraft.png}
%	\end{subfigure}
%	~
%	\begin{subfigure}{.45\textwidth}
%		\caption{Thanatological period fertility rates.}
%		\label{fig:FySpaghetti}
%		\includegraphics[scale=.45]{Figures/FySpaghettiDraft.png}
%	\end{subfigure}
%	}
%	\\
%	\end{center}
%	\begin{tiny}
%	*AUT, $1951-2010$; BGR, $1947-2009$; BLR, $1964-2012$; CAN, $1921-2009$; 
%	CHE, $1932-2011$; CHL, $1992-2005$; CZE, $1950-2011$; DEUTE, $1956-2011$; 
%	DEUTNP, $1990-2011$; DEUTW, $1956-2011$; ESP, $1922-2012$; EST, $1959-2011$; 
%	FIN, $1939-2009$; FRATNP, $1946-2013$; GBR\_NIR, $1974-2011$; GBR\_NP,
%	$1974-2011$; GBR\_SCO, $1945-2011$; GBRTENW, $1938-2011$; HUN, $1950-2009$;
%	IRL, $1955-2009$; ISL, $1963-2010$; JPN, $1947-2012$; LTU, $1959-2011$; NLD, $1950-2009$; 
%	NOR, $1967-2009$; PRT, $1940-2012$; RUS, $1959-2010$; SVK, $1950-2009$; 
%	SVN, $1983-2009$; SWE, $1891-2011$; TWN, $1976-2010$; UKR, $1959-2009$; 
%	USA, $1933-2010$
%	\end{tiny}
%\end{figure}
%
%In general, thanatological fertility rates will not be useful for purposes of
%projection, unless their empirical regularity is
%stronger than typical ASFR. This is a question I do not seek to answer
%in the present treatment, but in any case such rates are not susceptible to
%direct measurement.
%There are alternative ways of defining thanatological fertility rates, for instance by transforming the birth and
%population vector from a stable population. As proven in appendix~\ref{app:B},
%thanatological fertility rates derived from a stable population imply a certain
%symmetry with stable populations from the chronological model. 
%
\FloatBarrier
\section*{The years-left renewal model}
The renewal model that follows uses thanatological fertility rates, $\gamma(y)$,
the form of which is suggested in Figure~\ref{fig:FySpaghetti}, and we refer to
a homogenous, closed, unisex population based on female vital rates. Assuming constant vital rates, the births for the present year are given by 
\begin{align}
B(t) =& \int _0^\infty P(y,t)\gamma(y) \dd y = \int _0^\infty
P(a,t)m(a) \dd a \ec \intertext{where $a$ indexes chronological age, $y$ indexes
thanatological age, $P(a), P(y)$ are population counts, and $\gamma(y), m(a)$ are
exact fertility probabilities (rates), best imagined
as the fertility rates of females born to females. The thanatological
(first) integral can be broken down back in terms of chronological age:} =&
\int_{y=0}^\infty \int_{a=0}^\infty \gamma(y)P(a,t)f(y|a) \dd a \dd y \ec
\intertext{where $f(y|a)$ is defined per equation~\eqref{eq:vaupel1}, so
$d(a)/\ell(0)$.
We can relate the present population to past births with $P(a,t) = B(t-a)\ell(a)$:}
\label{eq:ergostep1}
=& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) B(t-a)d(a+y)\dd a \dd y \ep 
\intertext{Eventually strong ergodicity will assert itself, and $B(t)$ will
be related to $B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
intrinsic rate of growth:}
\label{eq:ergostep2}
=& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) B(t)e^{-ra}d(a+y)\dd a \dd y
\ep
\intertext{Divide out $B(t)$ to get back a familiar-looking renewal equation:}
\label{eq:thanoren}
1 =& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) d(a+y)e^{-ra}\dd a \dd y
\ep
\end{align}
Now compare this to Lotka's chronological formulation in
Equation~\eqref{eq:lotka},which can be expressed as well as:
\begin{equation}
 1 = \int_{a=0}^\infty \int_{y=0}^\infty m(a)d(a+y)e^{-ra}\dd y \dd a \ec
  \label{eq:lotka2}
\end{equation}
and note that \eqref{eq:lotka2} and \eqref{eq:thanoren} are really quite
similar, since $\int _{y=0}^\infty d(a+y)\dd a = l(a)$. For intuition, notice
that $l(a)$ is here decomposed into $d(a)$. Imagine a matrix of this
decomposition, where one margin is chronological age and the other margin is
thanatological age.
For the chronological case \eqref{eq:lotka}, we multiply chronological
age-specific fertility rates, $m(a)$, over the chronological age margin, and for the thanatological case we multiply $\gamma(y)$ over the thanatological age margin.
